PineAPPL:\\ combining EW and QCD corrections for PDF fits

### 2020-08-13 paper/sec-results.tex
section 3

In this section we demonstrate the capabilities of \textsc{PineAPPL} by
computing fast-in\-ter\-po\-la\-tion grids, accurate to NLO QCD and NLO QCD+EW,
for a representative set of processes in which EW corrections may have a
sizeable effect on the accuracy of the theoretical predictions.
In order to consider realistic kinematics for these
processes, we rely on the measurements that are commonly included in PDF
fits. Our aim is twofold. First, we want to validate the results
obtained with \textsc{PineAPPL}; second, we want to assess the
impact of the EW corrections for the specific experimental setups. We start by describing the
setup employed for the computation, then we discuss the features of each process, along with
the corresponding results.

@@

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc MG5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme~\cite{Denner:1999gp,Denner:2005fg} for all massive particles, see
ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that involve only massless particles in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: 
 in particular, PI contributions are consistently included at LO and NLO.\footnote{We employ the $\overline{G}_\mu$ scheme also for the QED coupling entering vertices involving initial-state photons, see section~4.3.3 of ref.\cite{Denner:2019vbn}.}
We use a PDF set that contains a photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast-interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation. Monte Carlo
weights are stored as Lagrange-interpolation grids.

The central values of the renormalisation and factorisation scales, $\mu_\mathrm{R}$ and
$\mu_\mathrm{F}$, are chosen in a process specific way, and are discussed in sec.~\ref{subsec:processes_and_measurements}. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted on-the-fly in the Monte Carlo generation upon scale
variations, with the technique presented in Ref.~\cite{Frederix:2011ss}. To this purpose, we use the default \textsc{MG5\_aMC}
implementation, whereby the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and the envelope
from the nine-point scale variations is constructed. However, we note that
\textsc{PineAPPL} allows the user to determine the envelope with any point
prescription, see appendix~\ref{app:pineappl-demo} for an example.

The values of the relevant physical parameters are chosen as
\begin{equation}
\begin{aligned}
M_\mathrm{W} &= \SI{80}{\giga\electronvolt} \text{,} \quad &
M_\mathrm{Z} &= \SI{91.176}{\giga\electronvolt} \text{,} \quad &
m_\mathrm{t} &= \SI{172.5}{\giga\electronvolt} \text{,} \quad \\
\Gamma_\mathrm{W} &= \SI{2.09}{\giga\electronvolt} \text{,} &
\Gamma_\mathrm{Z} &= \SI{2.50}{\giga\electronvolt} \text{,}
\end{aligned}
\label{eq:parameters}
\end{equation}
where $M_\mathrm{W}$, $M_\mathrm{Z}$ and $m_\mathrm{t}$ are the values of the W-boson, Z-boson, and
top-quark masses, respectively, and $\Gamma_\mathrm{W}$ and $\Gamma_\mathrm{Z}$ are the width of
the W boson and of the Z boson, respectively. The value of the strong
coupling is chosen consistently with the PDF set, $\alphas(M_\mathrm{Z})=0.118$.
The definition of observables and cuts is process-specific, and it follows the corresponding experimental measurements.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon distance. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematics observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination would have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections,
see refs.~\cite{Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order as the EW corrections, they are not included in our
computations. In fact, while nothing prevents to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of \enquote{standard} EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases {\bf MZ OTHERS???}).

@@

We focus on the following three processes: Drell--Yan lepton-pair production,
top-quark pair production, and Z-boson (lepton-pair) production with non-zero
transverse momentum in proton-proton collisions. For each of these processes,
we present the experimental measurements, process-specific settings, and phenomenological results.

When presenting the results, for each of the processes and measurements considered, we compute differential
cross sections for the observables defined in the experimental analyses in two different
ways: directly, by means of \textsc{MG5\_aMC}, and a posteriori, by convolving
the fast-interpolation grid produced by \textsc{PineAPPL} with the PDF set
specified in section~\ref{subsec:computational_settings}. We
will refer to the first result as the MC result, and to the second as
the \textsc{PineAPPL} result. We repeat the computation for theories accurate
to NLO QCD and to NLO QCD+EW, respectively. The corresponding orders of the
strong and electroweak couplings that we consider are specified for each process. In each case, we determine the
PDF uncertainty (coming from the PDF ensemble), the scale uncertainty (coming
from variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of \SI{0.1}{\permille} on the integrated cross section. While
this choice does not affect the validation of the \textsc{PineAPPL} result
against the MC result, it ensures that the statistical uncertainty of
the computation remains negligible in comparison to the PDF and scale
uncertainties, as we will explicitly demonstrate. This is a desirable feature
to correctly interpret the size of the EW corrections.

@@

\paragraph{Numerical results.}
We first consider the single-differential measurement of a lepton-pair
for large invariant masses performed by the ATLAS experiment at \SI{7}{\tera\electronvolt}.
From figure~\ref{fig:atlaszhighmass49fb} we immediately
observe that the validation of the \textsc{PineAPPL} result against the MC
result is successful. The relative difference between the two is of the order
\SI{0.1}{\permille} at most, with negligible fluctuations across different
invariant mass bins. The agreement is similarly good irrespective of the
perturbative accuracy of the theory (NLO QCD or NLO QCD+EW) or of the scale
choice.

@@

The measurement is mainly driven by $\mathrm{q}\bar{\mathrm{q}}$ scattering: specifically, the
leading (next-to-leading) contribution comes from a $\mathrm{u}\bar{\mathrm{u}}$/$\mathrm{c}\bar{\mathrm{c}}$ ($\mathrm{d}\bar{\mathrm{d}}$/$\mathrm{s}\bar{\mathrm{s}}$)
parton luminosity, which accounts for about \SI{55}{\percent} (\SI{49}{\percent}) of the cross section
for the lowest invariant mass bins, and \SI{68}{\percent} (\SI{22}{\percent}) for the largest invariant
mass bins.\footnote{The size of these contributions may depend on the input PDF
  set. Here and in the following, we always quote results obtained from the
  \texttt{NNPDF31\_nlo\_as\_0118\_luxqed} PDF set.}
The PI contribution raises from about \SI{1.3}{\percent} in the lowest bin to
about \SI{3.6}{\percent} in the highest bin. For the MRST2004qed PDF set, which is used to subtract PI contributions (see section~\ref{sec:doublecounting}), the value in the largest highest bin is roughly twice as large, \SI{6.9}{\percent} (also in absolute numbers).
Overall, the EW corrections range between \SI{5}{\percent}
around $M_{\ell\bar\ell}\sim \SI{150}{\giga\electronvolt}$, \SIrange{2}{3}{\percent} for intermediate invariant mass
values, $\SI{150}{\giga\electronvolt}\lesssim M_{\ell\bar\ell}\lesssim \SI{700}{\giga\electronvolt}$, and
\SIrange{15}{20}{\percent} for the largest invariant mass bin, $M_{\ell\bar\ell}>\SI{1000}{\giga\electronvolt}$,
see figure~\ref{fig:atlaszhighmass49fb}. For this reason, the data points with
$M_{\ell\bar\ell}>\SI{210}{\giga\electronvolt}$ were not included in the NNPDF3.1
analysis~\cite{Ball:2017nwa}.

@@

In comparison to the PDF uncertainty, the size of the EW corrections is
always larger, especially at the boundaries of the distribution. This fact
suggests that, once included in a global fit, EW corrections could recast the
relative weight of each data set included in the fit, and possibly lead to
an improvement in their overall description. In comparison to the scale
uncertainty, the size of the EW correction is similar, except for the four bins
at the largest invariant mass, where the latter is significantly larger than
the former. The size of the scale variation is the same for the QCD and the
QCD+EW theories. These facts suggest that the NNLO QCD correction is comparable
to the NLO QCD+EW correction, except at very large values of the invariant mass,
where the EW correction still dominates. This result stresses the need to
include the EW corrections in order to obtain an accurate description of the
large invariant mass bins. Finally, the Monte Carlo statistical
uncertainty remains negligible in comparison to the data, PDF and scale
uncertainties, and to the size of the EW correction. Our conclusions should
therefore not be affected by a generation of too few Monte Carlo events.

@@

%-------------------------------------------------------------------------------
\begin{figure}[!p]
    \centering
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin3}%
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin4}\\
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin5}%
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin6}\\
    \caption{Same as figure~\ref{fig:atlaszhighmass49fb}, but for the CMS
      double-differential Drell--Yan lepton pair measurement at a
      centre-of-mass energy of \SI{7}{\tera\electronvolt}~\cite{Chatrchyan:2013tia}. Displayed are
      only two of the six invariant mass bins available, respectively below the
      Z-boson mass peak, $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$, on the Z-boson mass
      peak, $\SI{60}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{120}{\giga\electronvolt}$, above the mass peak,
      $\SI{120}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{200}{\giga\electronvolt}$, and at very high invariant masses,
      $\SI{200}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$.
      Results for the slices $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ and $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ can be found in figure~\ref{fig:cmsdy2d11_bins12}.}
    \label{fig:cmsdy2d11_bins3456}
\end{figure}
%-------------------------------------------------------------------------------

@@

In general, the size of the EW corrections is comparable to or slightly larger
than the data uncertainty, except for the invariant mass bin 
$\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$,
where the shift due to the EW correction overshoots the data uncertainty by
about a factor of ten, and at large rapidities, where the shift due to the EW
correction, although it can become large, is always a fraction of the data
uncertainty. Because EW effects are subtracted from the data used
in PDF fits (see section~\ref{sec:doublecounting}), a good agreement between
data and theory is usually achieved without the inclusion of EW corrections.
However, as already observed in the case of the ATLAS measurement, should EW
corrections be included in a fit of PDFs, these are likely to become more
accurate: even though the apparent description of the data will not
improve, by including the more precisely predicted bins in the low invariant mass range, PDFs
will resemble more closely the underlying truth. Furthermore, the kinematic cut
that excludes any data point at large invariant mass and/or rapidity can be
safely removed.

In comparison to the PDF uncertainty, the size of the EW corrections is always
larger. We therefore anticipate that, even if the agreement between the more
accurate theory (including EW corrections) and the data will remain the same,
the PDFs will however become overall more accurate. In comparison to the scale
uncertainty, the size of the EW correction is similar, except on the Z-boson
mass peak, $\SI{60}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{120}{\giga\electronvolt}$, where the scale uncertainty exceeds
the size of the EW correction by about a factor of five. This is due to the choice
of invariant mass window around the Z peak, in which positive and negative EW corrections almost cancel.
Finally, the Monte Carlo statistical uncertainty remains negligible
in comparison to the data, PDF and scale uncertainties, and to the size of the
EW correction, except for a couple of bins at forward/backward rapidity in the highest
invariant mass bins. Improving the Monte Carlo precision will require to
generate a larger number of events, possibly with cuts that select
only the kinematic bins affected by the largest MC uncertainties. If this
turned out to be computationally too expensive, it would be desirable to treat
this uncertainty as an additional theoretical uncertainty in the PDF
fit~\cite{Ball:2018lag}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Top-quark pair production.}
\paragraph{Experimental measurements and process features.}
We select the single-differential distribution in either the transverse
momentum of the top quark, $p_\mathrm{T}^\mathrm{t}$, or the invariant mass of the top-quark
pair, $m_{\mathrm{t}\bar{\mathrm{t}}}$, measured by the ATLAS and CMS experiments at a centre-of-mass
energy of \SI{8}{\tera\electronvolt}~\cite{Aad:2015mbv,Khachatryan:2015oqa}. These measurements have
been extensively studied in the context of PDF fits in
refs.~\cite{Czakon:2016olj,Bailey:2019yze,Amoroso:2020lgh,Kadir:2020yml} (see
also refs.~\cite{Pagani:2016caq,Czakon:2017wor} for studies related to the photon density) and
included by default in the CT18~\cite{Hou:2019efy} analysis.
Because EW corrections are significantly smaller for distributions differential
in the rapidity of either the top quark or the top-quark
pair~\cite{Czakon:2017wor}, these distributions were preferred for inclusion
in the NNPDF3.1 analysis~\cite{Ball:2017nwa}. The process receives
pure QCD contributions at LO, $\mathcal{O}(\alphas^2)$, and
at NLO, $\mathcal{O}(\alphas^3)$. They make up our NLO QCD
computation. The NLO QCD+EW computation includes the $\mathcal{O}(\alphas^2)$
and $\mathcal{O}(\alphas^3)$ QCD contributions, the LO contribution
$\mathcal{O}(\alphas\alpha)$ and the NLO contribution
$\mathcal{O}(\alphas^2\alpha)$.
We do not consider the LO contribution $\mathcal{O}(\alpha^2)$ nor the
NLO contributions $\mathcal{O}(\alphas\alpha^2)$ and $\mathcal{O}(\alpha^3)$, which have been shown to be negligible~\cite{Czakon:2017wor,Frederix:2018nkq}.
EW corrections for this process
were computed in refs.~\cite{Bernreuther:2010ny,Hollik:2011ps,Kuhn:2011ri,Bernreuther:2012sx,Pagani:2016caq,Czakon:2017wor,Czakon:2017lgo,Czakon:2017mmr,Czakon:2019bcq,Czakon:2019txp}. The process receives contributions from
7 (37) parton luminosities at NLO QCD (NLO QCD+EW),
see appendix~\ref{app:lumis} for details.

@@

The process receives its leading contribution from $\mathrm{qg}$ and
$\bar{\mathrm{q}}g$ initiated channels, which account for about 65\%
of the cross section, with some variations in the relative contributions from
individual quark and atiquark flavours as the transverse momentum of the
$\mathrm{Z}$ boson varies; the PI contribution always remain negligible.
Overall, the EW corrections suppress the theoretical predictions by about
2\% -- 7\% as the transverse momentum of the $\mathrm{Z}$ boson increases.
The size of this shift is usually as large as the data uncertainty, and
exceeds it at large values of $p_\mathrm{T}^\mathrm{Z}$. As a consequence,
we anticipate the inclusion of EW corrections to be relevant for an accurate
fit to this data set.

The size of the EW correction is between four and six times larger than the
size of the PDF uncertainty: as previously noted in th eother cases, this fact
suggests that, once included in a PDF fit, EW corrections can improve the
accuracy of the PDFs. In comparison to the scale ucnertainty, the size of the
EW corrections remains negligible, as NNLO QCD corrections are comparatively
large, as expected in a process mostly initiated by gluons. They should
therefore be included along with EW corrections in a fit of PDFs. Finally, the
Monte Carlo uncertainty is well under control, as it remains mostly negligible
in comparison to the PDF, scale and data uncertainty.

>>>>

In this section we demonstrate the capabilities of \textsc{MG5\_aMC}+\textsc{PineAPPL} by
computing fast-interpolation grids, accurate to NLO QCD and NLO QCD+EW,
for a representative set of processes in which EW corrections may have a
sizeable effect on the accuracy of the theoretical predictions.
In order to consider some realistic kinematics for these
processes, we rely on a set of measurements that are commonly included in PDF
fits. Our aim is twofold. First, we want to validate the results
obtained with \textsc{PineAPPL}; second, we want to assess the
impact of the EW corrections for the specific experimental setups. We describe
the settings employed for the computations, and the corresponding results for
each process considered.

@@

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc MG5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme~\cite{Denner:1999gp,Denner:2005fg} for all massive particles, see
ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that involve only massless particles in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: 
 in particular, photon-induced (PI) contributions are consistently included at LO and NLO.\footnote{We employ the $\overline{G}_\mu$ scheme also for the QED coupling entering vertices involving initial-state photons, see section~4.3.3 of ref.\cite{Denner:2019vbn}.}
We use a PDF set that contains a photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast-interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation. Monte Carlo
weights are stored as Lagrange-interpolation grids.

The central values of the renormalisation and factorisation scales, $\mu_\mathrm{R}$ and
$\mu_\mathrm{F}$, are chosen in a process-specific way, as discussed in sec.~\ref{subsec:processes_and_measurements}. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted on-the-fly in the Monte Carlo generation upon scale
variations, with the technique presented in Ref.~\cite{Frederix:2011ss}. To this purpose, we use the default \textsc{MG5\_aMC}
implementation, whereby the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and the envelope
from the nine-point scale variations is constructed,
see equation~\eqref{eq:}. However, we note that
\textsc{PineAPPL} allows the user to determine the envelope with any point
prescription, see appendix~\ref{app:pineappl-demo} for an example.

The values of the relevant physical parameters are chosen as
\begin{equation}
%\begin{aligned}
%M_\mathrm{W} &= \SI{80}{\giga\electronvolt} \text{,} \quad &
M_\mathrm{Z} = \SI{91.176}{\giga\electronvolt} \text{,} \quad 
m_\mathrm{t} = \SI{172.5}{\giga\electronvolt} \text{,} \quad 
%\Gamma_\mathrm{W} &= \SI{2.09}{\giga\electronvolt} \text{,} &
\Gamma_\mathrm{Z} = \SI{2.50}{\giga\electronvolt} \text{,}
%\end{aligned}
\label{eq:parameters}
\end{equation}
where $M_\mathrm{Z}$ and $m_\mathrm{t}$ are the values of the Z-boson, and
top-quark masses, respectively, and $\Gamma_\mathrm{Z}$ is the width of
the Z boson. The value of the strong
coupling is chosen consistently with the PDF set, $\alphas(M_\mathrm{Z})=0.118$.

The definition of observables and cuts is process-specific, and it follows the corresponding experimental measurements,
see secion~\ref{sec:subsec:processes_and_measurements}.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon distance. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematics observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination would have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections,
see refs.~\cite{Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order of the EW corrections, they are not included in our
computations. In fact, while nothing prevents to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of \enquote{standard} EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases {\bf MZ OTHERS???}).

@@

We focus on the following three processes: Drell--Yan lepton-pair production,
top-quark pair production, and Z-boson (lepton-pair) production with non-zero
transverse momentum in proton-proton collisions. In the following, we shall
present the experimental measurements, the process-specific settings, and
the phenomenological results for each of these processes.

When presenting the results, for each of the processes and measurements considered, we compute differential
cross sections for the observables defined in the experimental analyses in two different
ways: directly, by means of \textsc{MG5\_aMC}, and a posteriori, by convolving
the fast-interpolation grid produced by \textsc{PineAPPL} with the PDF set
specified in section~\ref{subsec:computational_settings}. We
refer to the first result as the MC result, and to the second as
the \textsc{PineAPPL} result. We repeat the computation for theories accurate
to NLO QCD and to NLO QCD+EW, respectively. The corresponding orders of the
strong and EW couplings that we consider are specified for each process. In each case, we determine the
PDF uncertainty (coming from the PDF ensemble), the scale uncertainty (coming
from variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of \SI{0.1}{\permille} on the integrated cross section. While
this choice does not affect the validation of the \textsc{PineAPPL} result
against the MC result, it ensures that the statistical uncertainty of
the computation remains negligible in comparison to the PDF and scale
uncertainties, as we will explicitly demonstrate. This is a desirable feature
to correctly interpret the size of the EW corrections.

@@

\paragraph{Numerical results.}
We first consider the single-differential measurement of a lepton-pair
for large invariant masses performed by the ATLAS experiment at \SI{7}{\tera\electronvolt}.
From figure~\ref{fig:atlaszhighmass49fb} we immediately
observe that the validation of the \textsc{PineAPPL} result against the MC
result is successful. The relative difference between the two is of the order of
\SI{0.1}{\permille} at most, with negligible fluctuations across different
invariant mass bins. The agreement is similarly good irrespective of the
perturbative accuracy of the theory (NLO QCD or NLO QCD+EW) or of the scale
choice.

@@

The measurement is mainly driven by $\mathrm{q}\bar{\mathrm{q}}$ scattering: specifically, the
leading (next-to-leading) contribution comes from a $\mathrm{u}\bar{\mathrm{u}}$/$\mathrm{c}\bar{\mathrm{c}}$ ($\mathrm{d}\bar{\mathrm{d}}$/$\mathrm{s}\bar{\mathrm{s}}$)
parton luminosity, which accounts for about \SI{55}{\percent} (\SI{49}{\percent}) of the cross section
for the lowest invariant mass bins, and \SI{68}{\percent} (\SI{22}{\percent}) for the largest invariant
mass bins.\footnote{The size of these contributions may depend on the input PDF
  set. Here and in the following, we always quote results obtained from the
  \texttt{NNPDF31\_nlo\_as\_0118\_luxqed} PDF set.}
The PI contribution raises from about \SI{1.3}{\percent} in the lowest bin to
about \SI{3.6}{\percent} in the highest bin.\footnote{For the \texttt{MRST2004qed} PDF set~\cite{Martin:2004dh}, which is used to subtract PI contributions (see section~\ref{sec:doublecounting}), the value in the highest bin is roughly twice as large, \SI{6.9}{\percent} (also in absolute numbers).}
Overall, the EW corrections range between \SI{5}{\percent}
around $M_{\ell\bar\ell}\sim \SI{150}{\giga\electronvolt}$, \SIrange{2}{3}{\percent} for intermediate invariant mass
values, $\SI{150}{\giga\electronvolt}\lesssim M_{\ell\bar\ell}\lesssim \SI{700}{\giga\electronvolt}$, and
\SIrange{15}{20}{\percent} for the largest invariant mass bin, $M_{\ell\bar\ell}>\SI{1000}{\giga\electronvolt}$,
see figure~\ref{fig:atlaszhighmass49fb}. For this reason, the data points with
$M_{\ell\bar\ell}>\SI{210}{\giga\electronvolt}$ were not included in the NNPDF3.1
analysis~\cite{Ball:2017nwa}.

@@

In comparison to the PDF uncertainty, the size of the EW corrections is
always larger, especially at the boundaries of the distribution. This fact
suggests that, once included in a global fit, EW corrections will make PDFs
more accurate. In comparison to the scale
uncertainty, the size of the EW correction is similar, except for the four bins
at the largest invariant mass, where the latter is significantly larger than
the former. This fact suggests that the NNLO QCD correction is comparable
to the NLO QCD+EW correction, except at very large values of the invariant mass,
where the EW correction still dominates. This result stresses the need to
include the EW corrections in order to obtain an accurate description of the
large invariant mass bins. Finally, the Monte Carlo 
uncertainty remains negligible in comparison to the data, PDF and scale
uncertainties, and to the size of the EW correction. Our conclusions should
therefore not be affected by a generation of too few Monte Carlo events.

@@

%-------------------------------------------------------------------------------
\begin{figure}[!p]
    \centering
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin3}%
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin4}\\
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin5}%
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin6}\\
    \caption{Same as figure~\ref{fig:atlaszhighmass49fb}, but for the CMS
      double-differential Drell--Yan lepton pair measurement at a
      centre-of-mass energy of \SI{7}{\tera\electronvolt}~\cite{Chatrchyan:2013tia}. Displayed are
      only four of the six invariant mass bins available, respectively below the
      Z-boson mass peak, $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$, on the Z-boson mass
      peak, $\SI{60}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{120}{\giga\electronvolt}$, above the mass peak,
      $\SI{120}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{200}{\giga\electronvolt}$, and at very high invariant masses,
      $\SI{200}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$.
      Results for the slices $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ and $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ can be found in figure~\ref{fig:cmsdy2d11_bins12}.}
    \label{fig:cmsdy2d11_bins3456}
\end{figure}
%-------------------------------------------------------------------------------

@@

In general, the size of the EW corrections is comparable to or slightly larger
than the data uncertainty, except for the invariant mass bin 
$\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$,
where the shift due to the EW correction overshoots the data uncertainty by
about a factor of ten, and at large rapidities, where the shift due to the EW
correction, although it can become large, is always a fraction of the data
uncertainty. Because EW effects are subtracted from the data used
in PDF fits (see section~\ref{sec:doublecounting}), a good agreement between
data and theory is usually achieved without the inclusion of EW corrections.
However, as already observed in the case of the ATLAS measurement, should EW
corrections be included in a fit of PDFs, the latter are likely to become more
accurate: even though the apparent description of the data will not
improve, by including the more precisely predicted bins in the low invariant mass range, PDFs
will resemble more closely the underlying truth. Furthermore, the kinematic cut
that excludes any data point at large invariant mass and/or rapidity can be
safely removed.

In comparison to the PDF uncertainty, the size of the EW corrections is always
larger. We therefore anticipate that, even if the agreement between the more
accurate theory (including EW corrections) and the data will remain the same,
the PDFs will however become overall more accurate. In comparison to the scale
uncertainty, the size of the EW correction is similar, except on the Z-boson
mass peak, $\SI{60}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{120}{\giga\electronvolt}$, where the scale uncertainty exceeds
the size of the EW correction by about a factor of five. This is due to the choice
of invariant mass window around the Z peak, in which positive and negative EW corrections almost cancel.
Finally, the Monte Carlo uncertainty remains negligible
in comparison to the data, PDF and scale uncertainties, and to the size of the
EW correction, except for a couple of bins at forward/backward rapidity in the highest
invariant mass bins. Improving the Monte Carlo precision will require to
generate a larger number of events, possibly with cuts that select
only the kinematic bins affected by the largest MC uncertainties. If this
turned out to be computationally too expensive, it would be desirable to treat
this uncertainty as an additional theoretical uncertainty in the PDF
fit~\cite{Ball:2018lag}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Top-quark pair production.}
\paragraph{Experimental measurements and process features.}
We select the single-differential distribution in either the transverse
momentum of the top quark, $p_\mathrm{T}^\mathrm{t}$, or the invariant mass of the top-quark
pair, $m_{\mathrm{t}\bar{\mathrm{t}}}$, measured by the ATLAS and CMS experiments at a centre-of-mass
energy of \SI{8}{\tera\electronvolt}~\cite{Aad:2015mbv,Khachatryan:2015oqa}. These measurements have
been extensively studied in the context of PDF fits in
refs.~\cite{Czakon:2016olj,Bailey:2019yze,Amoroso:2020lgh,Kadir:2020yml} (see
also refs.~\cite{Pagani:2016caq,Czakon:2017wor} for studies related to the photon density) and
included by default in the CT18~\cite{Hou:2019efy} analysis.
Because EW corrections are significantly smaller for distributions differential
in the rapidity of either the top quark or the top-quark
pair~\cite{Czakon:2017wor}, these distributions were preferred for inclusion
in the NNPDF3.1 analysis~\cite{Ball:2017nwa}. The process receives
pure QCD contributions at LO, $\mathcal{O}(\alphas^2)$, and
at NLO, $\mathcal{O}(\alphas^3)$. These orders make up our NLO QCD
computation. The NLO QCD+EW computation includes the $\mathcal{O}(\alphas^2)$
and $\mathcal{O}(\alphas^3)$ QCD contributions, the LO contribution
$\mathcal{O}(\alphas\alpha)$ and the NLO contribution
$\mathcal{O}(\alphas^2\alpha)$.
We do not consider the LO contribution $\mathcal{O}(\alpha^2)$ nor the
NLO contributions $\mathcal{O}(\alphas\alpha^2)$ and $\mathcal{O}(\alpha^3)$, which have been shown to be negligible~\cite{Czakon:2017wor,Frederix:2018nkq}.
EW corrections for this process
were computed in refs.~\cite{Bernreuther:2010ny,Hollik:2011ps,Kuhn:2011ri,Bernreuther:2012sx,Pagani:2016caq,Czakon:2017wor,Czakon:2017lgo,Czakon:2017mmr,Czakon:2019bcq,Czakon:2019txp}. The process receives contributions from
7 (37) parton luminosities at NLO QCD (NLO QCD+EW),
see appendix~\ref{app:lumis} for details.

@@

The process receives its leading contribution from $\mathrm{qg}$- and
$\bar{\mathrm{q}}\mathrm{g}$-initiated channels, which account for about 65\%
of the cross section, with some variations in the relative contributions from
individual quark and atiquark flavours as the transverse momentum of the
$\mathrm{Z}$ boson varies; the PI contribution always remain negligible.
Overall, the EW corrections suppress the theoretical predictions by about
2\% -- 7\% as the transverse momentum of the $\mathrm{Z}$ boson increases.
The size of this shift is usually as large as the data uncertainty, and
exceeds it at large values of $p_\mathrm{T}^\mathrm{Z}$. As a consequence,
we anticipate the inclusion of EW corrections to be relevant for an accurate
fit to this data set.

The size of the EW correction is between four and six times larger than the
size of the PDF uncertainty: as previously noted in the other cases, this fact
suggests that, once included in a PDF fit, EW corrections can improve the
accuracy of the PDFs. In comparison to the scale uncertainty, the size of the
EW corrections remains negligible, as NNLO QCD corrections are comparatively
large, as expected in a process mostly initiated by gluons. They should
therefore be included along with EW corrections in a fit of PDFs. Finally, the
Monte Carlo uncertainty is well under control, as it remains mostly negligible
in comparison to the PDF, scale and data uncertainty, and to the size of the
EW correction.
